---
title: "simpleRegress"
author: "Michael Maida"
date: "Fall 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
bears$MONTH <- as.factor(bears$MONTH)
bears$SEX <- as.factor(bears$SEX)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

## Model LENGTH ~ WEIGHT  this is an example  
```{r}
simp <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp)
summary.lm(simp)
```

The residual R^2 for the neck was 5.376, and the residual R^2 for the statistic using weight was 5.432. Additionally, the adjusted R^2 for the neck was 0.7476 while the adjusted R^2 for the weight was 0.7423. Based on those two factors, the neck proved to be a better model than the weight. We will have to try again.


## Model LENGTH ~ CHEST  this is an example  

```{r}
simp <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

While the residual standard error of the neck data was 5.376, the residual standard error of the newly calculated chest data was 4.939. Additionally, the adjusted R^2 of the neck data was 0.7476 while the newly calculated adjusted R^2 of the chest data was 0.7869. Since the residual standard error of the chest data was a smaller value while the adjusted R^2 value of the chest was larger, we know that the chest is a more accurate determiner of bear length. 

## Model LENGTH ~ AGE  this is an example  
```{r}
simp <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp)
summary.lm(simp)
```

## Model LENGTH ~ HEADLEN  this is an example  
```{r}
simp <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```

## Model LENGTH ~ HEADWTH  this is an example  
```{r}
simp <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```